Quantum H theorem, variational entropy, and relaxation processes

### 2020-12-21 tex/classical.tex
Rewording: up tp sec. 2.3.1

%-------------------------
\subsection{H-theorem and the Maxwell-Boltzmann distribution}

The evolution of a dilute gas towards thermodynamic equilibrium is frequently
addressed by first defining the functional \cite{bib:tolman,bib:huang}
%
\begin{equation}\label{eq:hfunctional}
   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
\end{equation} 
%
Here, $f(\vec{v},t)$ is a homogeneous distribution function, \textit{i.e.}
$f$ is assumed to be independent of $\vec r$.
The functional $H$, originally introduced by Boltzmann in 1872,
describes a diluted gas, at temperature $T$, with a total energy $E$, a total
number of free classical particles $N$, and occupying a volume $V$.

Subsequently, the time-evolution of the gas towards equilibrium is obtained through the
$H$-theorem. This theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
transport equation and the molecular chaos hypothesis is valid, then
$dH/dt\leq0$ in general, and if $dH/dt=0$, then the system is in the equilibrium state.
The $H$-theorem is proven by taking the time derivative
of Eq. (\ref{eq:hfunctional}) and using the Boltzmann transport equation
(\ref{eq:transport}) \cite{bib:huang}. The $H$-theorem fixes the direction of time-evolution
of systems in nature, which is consistent with the second law of thermodynamics; in fact,
$H$ can be associated with an entropy density.

@@

%---------------------------
\subsection{Out of equilibrium, non-homogeneous distributions}

As we saw in the previous section,
the validity of the $H$-theorem relies significantly on assuming
that the distribution function is homogeneous and that the hypothesis of
molecular chaos is surmised. However, in order to consider systems whose
distribution functions are not necessarily homogeneous, which
allows for studying systems out of equilibrium,
we introduce a modified $H$ functional as follows.

The proposed functional, denoted as $H'$, describes a diluted gas occupying a volume, $V$,
divided into $K$ cells, which, without loss of generality,
have local identical volumes (\textit{i.e.} $\delta V_M = V/K,\ M=1,\,\dots\,,K$).
Each cell has its own local
$H$ functional, $H'_M$, and local homogeneous distribution function,
$f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
local temperature, $T_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
global energy $E$, and a global number of free
classical particles $N$. The functional takes the form
%
\begin{equation}\label{eq:CH2}
   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%

The distribution functions, $\{f_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
relative to the global system, and on the velocity
$\vec{v}$ and time $t$. In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
%
\begin{subequations}\label{eq:cellrestrictions}
\begin{eqnarray}
    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
      \mathrm{d}^{3}v \label{Hcell},\\
    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\\
    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
\end{eqnarray}
\end{subequations}
%
Notice that assuming local homogeneous distribution functions means that we are surmising the 
local equilibrium hypothesis. In addition, $f_{M}(\vec{v},t)$ must satisfy the following
microcanonical restrictions:
%
\begin{subequations}\label{eq:micro}
\begin{equation}\label{eq:micron}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N
\end{equation}
and
\begin{equation}\label{eq:microe}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
\end{equation}
\end{subequations}
%

If the complete system is in equilibrium, then the local number
of particles and the local
energy do not depend on the cell number, this is
%
\begin{subequations}\label{eq:eqNandE}
\begin{equation}
   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
\end{equation}
%
and
%
\begin{equation}
   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
\end{equation}
\end{subequations}
%
In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.

We now use the variational method in order to maximize $H'$, applying the restrictions
given in Eqs.~(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) through Lagrange
multipliers ($\{\alpha_M\}$ and $\{\beta_M\}$). This yields
%
\begin{eqnarray}
    \frac{\delta H'}{\delta f_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
      \frac{\delta}{\delta f_J(\vec{v}')}\left[
        f_M(\vec{v})\ln f_M(\vec{v})
        \right]
       \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
       	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v\nonumber\\
    & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
    	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v \nonumber\\
    & = & \ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
\end{eqnarray}
%
Solving the last line for $f_J(\vec{v}')$ renders
%
\begin{comment}
\begin{eqnarray}
    \ln f_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
       -1\quad\Rightarrow\quad f_J(\vec{v}')\\
    & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{eq:relacion1},
\end{eqnarray}
\end{comment}
\begin{equation}\label{eq:relacion1}
	f_J(\vec{v}') = Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') }
\end{equation}
%
where $C$ is a constant. We notice that using the variational procedure on
$H'$ predicts that the
distribution function of each cell has the form of the Maxwell-Boltzmann
distribution function.

As $H'$ in equilibrium is proportional to entropy, it is trivial to think that
Lagrange multipliers do not depend on the position because of the system is in
equilibrium and consequently, the distribution function is homogeneous, that is
%
\begin{equation}
    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).
\end{equation} 
%
The constant $C$ can be omitted defining the following $H$ functional
%
\begin{equation}
   H''(t)=\sum_{M=1}^{K}\int \left[f_M(\vec{v},t)
    \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
\end{equation}
%
but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the
total particle number. As $N$ is a constant and any $H$ functional has sense
when we calculate its time derivative, this constant can be omitted.

On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{eq:CH2}) as
%
\begin{equation}
    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 
\end{equation}
%
where
%
\begin{equation}
    H_{boltz}(t)=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
\end{equation}
%
Besides, if we analyze the following expression
%
\begin{equation}
    \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}
\end{equation}
%
we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that
%
\begin{equation}
    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 
\end{equation}
%
We obtained the additive property when we supposed the distribution function is
spatial-homogeneous.

>>>>

%-------------------------
\subsection{A brief summary of the H-theorem and the Maxwell-Boltzmann distribution}

The evolution of a dilute gas towards thermodynamic equilibrium is frequently
addressed by first defining the functional \cite{bib:tolman,bib:huang}
%
\begin{equation}\label{eq:hbfunctional}
   H_{B}=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.
\end{equation} 
%
Here, $f(\vec{v},t)$ is a homogeneous distribution function, \textit{i.e.}
$f$ is assumed to be independent of $\vec r$.
The functional $H_B$,
originally introduced by Boltzmann in 1872,
describes a diluted gas, at temperature $T$, with a total energy $E$, a total
number of free classical particles $N$, and occupying a volume $V$.
(We deviate from the custommary notation in order to
clearly distinguish the Boltzmann functional $H$ from our proposals, and
$f_B$ to denote the Maxwell-Boltzmann distribution function.)

Under this scheme, the time-evolution of the gas towards equilibrium is obtained through the
$H$-theorem. This theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann
transport equation and the molecular chaos hypothesis is valid, then
$dH_B/dt\leq0$ in general, and if $dH_B/dt=0$, then the system is in the equilibrium state.
The $H$-theorem is proven by taking the time derivative
of Eq. (\ref{eq:hbfunctional}) and using the Boltzmann transport equation
(\ref{eq:transport}) \cite{bib:huang}. The $H$-theorem fixes the direction of time-evolution
of systems in nature, which is consistent with the second law of thermodynamics; in fact,
$H_B$ can be associated with an entropy density.

@@

%---------------------------
\subsection{Alternative classical $H$-functional}

As we saw in the previous section,
the validity of the $H$-theorem relies significantly on assuming
that the distribution function is homogeneous and that the hypothesis of
molecular chaos is surmised. However, in order to consider systems whose
distribution functions are not necessarily homogeneous, which
allows for studying systems out of equilibrium,
we introduce a modified $H$-functional as follows. Hereafter, we will
use the term homogeneous distribution function to refer to a distribution
function thad does not depend on the coordinates.

The proposed functional, denoted as $H'$, describes a diluted classical
gas occupying a volume, $V$,
divided into $K$ cells, which, without loss of generality,
have local identical volumes (\textit{i.e.} $\delta V_M = V/K,\ M=1,\,\dots\,,K$).
Each cell has its own local
$H$-functional, $H'_M$, and local homogeneous distribution function,
$f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
local temperature, $T_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
global energy $E$, and a global number of free
classical particles $N$, and it is considered to be perfectly isolated (\textit{i.e.} the
global system is a microcanonical ensemble). We also assume that the number of particles
in each cell is sufficiently large, so as to obtain accurate averages.
The functional takes the form
%
\begin{equation}\label{eq:CH2}
   H'(t)=\sum_{M=1}^{K}\int_{\delta V_M} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%

The distribution functions, $\{f_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
relative to the global system, and on the velocity
$\vec{v}$ and time $t$. Each $f_M$ can be formally extended to the complete coordinate space
by defining each $f_M$ to be zero outside the $M$-th cell, in such a manner that
the distribution function of the complete system is a piece-wise sum
of $\{f_M(\vec v,t)\}$:
%
\begin{equation}
   f(\vec r,\vec v,t)=\sum_{M=1}^Kf(\vec r_M,\vec v,t)=\sum_{M=1}^Kf_M(\vec v,t).
\end{equation}
%
Here $\vec r_M$ is the center of the cell of index $M$. This extended definition
allows us to omit the symbol $\delta V_M$ on all integrals performed over the cell volume.
In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
%
\begin{subequations}\label{eq:cellrestrictions}
\begin{eqnarray}
    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
      \mathrm{d}^{3}v \label{eq:Hcell},\\
    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\label{eq:Ncell}\\
    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v\label{eq:Ecell}.
\end{eqnarray}
\end{subequations}
%
Notice that assuming homogeneous distribution functions per cell means that we are surmising the 
local equilibrium hypothesis. In addition, the set $\{f_{M}(\vec{v},t)\}$ must satisfy the following
microcanonical restrictions:
%
\begin{subequations}\label{eq:micro}
\begin{equation}\label{eq:micron}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N
\end{equation}
and
\begin{equation}\label{eq:microe}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
\end{equation}
\end{subequations}
%

We now use the variational method in order to maximize $H'$, applying the restrictions
given in Eqs.~(\ref{eq:cellrestrictions}) and (\ref{eq:micro}) through Lagrange
multipliers ($\{\alpha_M\}$ and $\{\beta_M\}$). This yields
%
\begin{eqnarray}
    \frac{\delta H'}{\delta f_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
      \frac{\delta}{\delta f_J(\vec{v}')}\left[
        f_M(\vec{v})\ln f_M(\vec{v})
        \right]
       \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
       	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v\nonumber\\
    & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
    	\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}
      \mathrm{d}^3v \nonumber\\
    & = & \ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
\end{eqnarray}
%
Solving the last line for $f_J(\vec{v}')$ renders
%
\begin{comment}
\begin{eqnarray}
    \ln f_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
       -1\quad\Rightarrow\quad f_J(\vec{v}')\\
    & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{eq:relacion1},
\end{eqnarray}
\end{comment}
\begin{equation}\label{eq:relacion1}
	f_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
\end{equation}
%
where $C$ is a constant. We notice that using the variational procedure on
$H'$ predicts that the
distribution function of each cell has the form of the Maxwell-Boltzmann
distribution function, which is consistent with the local equilibrium assumption.

\subsubsection{Properties of $H'$ for systems in equilibrium}

If the complete system is in equilibrium without external forces applied to
the gas, from classical thermodynamics of
systems in equilibrium, we ascertain that the local number
of particles and the local
energy do not depend on the cell number, this is
%
\begin{subequations}\label{eq:eqNandE}
\begin{equation}
   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
\end{equation}
%
and
%
\begin{equation}
   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
\end{equation}
\end{subequations}
%
In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
Moreover, the distribution function is homogeneous, hence $f_M$
does not depend on the cell number $M$ (\textit{i.e.} $f_M(\vec v,t)=f(\vec v,t)$,
$\forall M$). Several properties follow directly from this, \textit{e.g.}
from Eqs.~(\ref{eq:eqNandE}) $E=\sum_M\mathcal{E}'_M=K\bar{\mathcal{E}}'$, $N=\sum_M\mathcal{N}'_M=K\bar{\mathcal{N}}'$ (here we have used Eqs. (\ref{eq:Ecell})
and (\ref{eq:Ncell})),
and Eq.~(\ref{eq:CH2}), in terms of Eq.~(\ref{eq:hbfunctional}), reduces to
%
\begin{equation}
    H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v
         = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{B}(t).
\end{equation}
%
In order to identify $H'$ with the entropy, it remains to show that $H'(t)$ is extensive,
with respect to $f'(\vec v,t)\big(\equiv Kf_(\vec v,t)\big)$.
This is shown by analyzing the following expression
%
\begin{eqnarray}\label{eq:sum-h}
    \int Kf(\vec{v},t) \ln \big[Kf(\vec{v},t)\big] \mathrm{d}^{3}v & = & 
        \int \big[(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)\big]\mathrm{d}^{3}v\nonumber\\
        & = & K\int\Big\{ f(\vec{v},t)\big[\ln K + \ln f(\vec{v},t)\big] \Big\}\mathrm{d}^{3}v.
\end{eqnarray}
%
We observe that if the number of particles in the $\mu$-space, $f(\vec{v},t)$, is much greater than
the number of cells $K$, then the first term of Eq.~(\ref{eq:sum-h}) is negligible, and consequently:
%
\begin{equation}\label{aditive-property}
    \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v \simeq
    K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,
\end{equation}
%
\textit{i.e.} $H_M'$ is extensive, and the sum $\sum_MH'_M$ is the $H$-functional of the complete
system and reduces to the Boltzmann $H$-functional. Hence, $H'$ can also be identified with the
system's entropy.

Furthermore, in equilibrium the Lagrange multipliers are position- and time-independent,
thus $f_M(\vec{v})$
reduces to
%
\begin{equation}
    f_M(\vec{v})=C\exp\left(\alpha+\beta \epsilon(\vec{v})\right)
    \equiv \bar{f}(\vec{v}),\quad M=1,\,\dots\,,K.
\end{equation} 
%
The constant $C$ can be omitted, which is shown by defining the following $H''$
functional:
%
\begin{equation}
   H''(t)=\sum_{M=1}^{K}\int \left[f_M(\vec{v},t)
    \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},
\end{equation}
%
however $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$.
Since $N$ is a constant and we are mainly intersted in the time-derivative of $H''$,
$C$ can be conveniently omitted. In other words, $H'$ renders the Maxwell-Boltzmann
distribution function for systems in equilibrium.

\subsubsection{Alternative $H$-theorem}

### 2020-12-21 tex/conclusions.tex
Typos, 1st intro paragraph, secs. 2.0, 2.1
Rewording: up tp sec. 2.3.1

\section{Conclusions}
We conclude the following statements:\\
\begin{enumerate}
    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
    \item We can obtain the transport equation from the $H$ functional.
    \item The variational entropy is proportional to the Boltzmann's $H$ functional in the limit of the high energy levels.
    \item We find an expression for the entropy to quantum gases.
    \end{enumerate}

>>>>

\section{Conclusions}\label{sec:conclusions}
We conclude the following statements:\\
\begin{enumerate}
    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
    \item We can obtain the transport equation from the $H$-functional.
    \item The variational entropy is proportional to the Boltzmann's $H$-functional in the limit of the high energy levels.
    \item We find an expression for the entropy to quantum gases.
    \end{enumerate}

### 2020-12-21 tex/disscussion.tex
Rewording: start of section 2.3
Rewording: up tp sec. 2.3.1

@@

In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis. However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume. Moreover, the Boltzmann's $H$-theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds. We can include the fact that we proved the Boltzmann's $H$-theorem without using the Boltzmann's transport equation. It was only necessary to specify the behavior of the deviation $g$ and its derivative in time. Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$ functional through the variational method. Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$ functional is not zero when the system is out of equilibrium. This is a consequence of the $H$-theorem.\\

@@

We also find that variational entropy is proportional to the Boltzmann's $H$ functional in the limit of high energy levels. This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}). Moreover, if we see the expression for the Boltzmann's $H$ functional in equilibrium

@@

    H_{Boltz}=-\frac{S}{Vk},

>>>>

%--------------------------------------
% Me parece que sera necesario incluir la discusin sobre los procesos de relajacin.
%Sin esa parte el contenido original del artculo podria parecer insuficiente para
%una buena publicacion%

@@

In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis. However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume. Moreover, the Boltzmann's $H$-theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds. We can include the fact that we proved the Boltzmann's $H$-theorem without using the Boltzmann's transport equation. It was only necessary to specify the behavior of the deviation $g$ and its derivative in time. Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$-functional through the variational method. Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$-functional is not zero when the system is out of equilibrium. This is a consequence of the $H$-theorem.\\

@@

We also find that variational entropy is proportional to the Boltzmann's $H$-functional in the limit of high energy levels. This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}). Moreover, if we see the expression for the Boltzmann's $H$-functional in equilibrium

@@

    H_{B}=-\frac{S}{Vk},

### 2020-12-21 tex/introduction.tex
Rewording: up tp sec. 2.3.1

In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
was described by Boltzmann through the Boltzmann transport equation.
Furthermore, the evolution of an equilibrium state for any system was proved by
means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
spatial-homogeneous distribution function hypothesis.
However, there are some aspects in the classical $H$-theorem and the
Boltzmann's $H$ functional that deserve some review to accomplish better
consistency, for example, the modification of the $H$-theorem in stochastic
trajectories, violations in the second law of thermodynamics, and the relation
between Shannon's measure of information and the Boltzmann's entropy %[6-10].
\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.

@@

On the other hand, some works in the quantum operator formalism attempt to
describe quantum transport phenomena using the Hamiltonian of the system and
the master equation (analogous to the Boltzmann transport equation)
\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. %[12-15]
Besides, some authors construct the $H$ functional and develop the proof of the
quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
However, the homogeneous distribution function hypothesis and the
correspondence principle are uncleared or not discussed. Moreover, the validity
of the quantum $H$-theorem and the second law of thermodynamics, and the
interpretation of the quantum entropy is still in discussion %[19-26].
\cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:acharya2019,%
bib:kastner2017,bib:gring2012,bib:han2015,bib:das2018}.

@@

The organization of this article is as follows.
In Section~\ref{sec:classical}, we highlight, to our purposes, the most
fundamental assumptions required to proof the Boltzmann $H$-theorem,
we provide an alternative method to obtain the Maxwell-Boltzmann
distribution using the variational method, and an alternative demonstration of
the $H$-theorem for classical systems. In Section~\ref{sec:quantum}
we review the Tolman proposal for the quantum
version of the $H$-theorem (quantum $H$-theorem)
\commr{and how the Bose-Einstein and Fermi-Dirac distributions are treated within
this framework},
and our alternative proof of the quantum $H$-theorem. In 
Section~\ref{sec:qccorrespondence}
the classical-quantum correspondence between the
quantum $H$ functional and the classical $H$ functional is analyzed.
In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
in a quantum ideal gas; here we provide a time-evolution equation that could be
considered the analogous of the classical Boltzmann transport equation. Finally,
we discuss some key ideas derived from our approach in Section~\ref{sec:disscussion},
and provide some conclusions in Section~\ref{sec:conclusions}.

>>>>

In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
was described by Boltzmann through the Boltzmann transport equation.
Furthermore, the evolution of an equilibrium state for any system was proved by
means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
spatial-homogeneous distribution function hypothesis.
However, there are some aspects in the classical $H$-theorem and the
Boltzmann's $H$-functional that deserve some review to accomplish better
consistency, for example, the modification of the $H$-theorem in stochastic
trajectories, violations in the second law of thermodynamics, and the relation
between Shannon's measure of information and the Boltzmann's entropy %[6-10].
\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.

@@

On the other hand, some works in the quantum operator formalism attempt to
describe quantum transport phenomena using the Hamiltonian of the system and
the master equation (analogous to the Boltzmann transport equation)
\cite{bib:gorban2014,bib:bennaim2017,bib:tolman,bib:li2019}. %[12-15]
Besides, some authors construct the $H$-functional and develop the proof of the
quantum $H$-theorem \cite{bib:silva2010,bib:deroeck2006,bib:grabert1974}.
However, the homogeneous distribution function hypothesis and the
correspondence principle are uncleared or not discussed. Moreover, the validity
of the quantum $H$-theorem and the second law of thermodynamics, and the
interpretation of the quantum entropy is still in discussion %[19-26].
\cite{bib:silva2010,bib:deroeck2006,bib:grabert1974,bib:acharya2019,%
bib:kastner2017,bib:gring2012,bib:han2015,bib:das2018}.

@@

The organization of this article is as follows.
In Section~\ref{sec:classical}, we highlight, to our purposes, the most
fundamental assumptions required to proof the Boltzmann $H$-theorem,
we provide an alternative method to obtain the Maxwell-Boltzmann
distribution using the variational method, and an alternative demonstration of
the $H$-theorem for classical systems. In Section~\ref{sec:quantum}
we review the Tolman proposal for the quantum
version of the $H$-theorem (quantum $H$-theorem)
\commr{and how the Bose-Einstein and Fermi-Dirac distributions are treated within
this framework},
and our alternative proof of the quantum $H$-theorem. In 
Section~\ref{sec:qccorrespondence}
the classical-quantum correspondence between the
quantum $H$-functional and the classical $H$-functional is analyzed.
In Section~\ref{sec:relaxproc} we explore how relaxation processes occur
in a quantum ideal gas; here we provide a time-evolution equation that could be
considered the analogous of the classical Boltzmann transport equation. Finally,
we discuss some key ideas derived from our approach in Section~\ref{sec:disscussion},
and provide some conclusions in Section~\ref{sec:conclusions}.

### 2020-12-21 tex/qccorrespondence.tex
quantum.tex reformatted: 80chars or less
Typos, 1st intro paragraph, secs. 2.0, 2.1
Rewording: start of section 2.3
Rewording: up tp sec. 2.3.1

\section{Quantum Classical correspondence}
%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$ functional in the appropriated limit.\\

@@

    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{Boltz},\label{h-quantic4}

@@

where $H_{Boltz}$ now is (\ref{CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{tolman} eq. (102.6) for more information.}\\

@@

that corresponds to the classical $H$ functional in equilibrium described in the energy space.}\\

>>>>

\section{Quantum Classical correspondence}\label{sec:qccorrespondence}
%The variational entropy $H$ that was defined in (\ref{entropy2}) can be reduce to the Boltzmann's $H$-functional in the appropriated limit.\\

@@

    &=& V \sum_n \left[N_n \ln N_n - N_n \ln \delta \epsilon_n  \right]=\de V_M K H_{B},\label{h-quantic4}

@@

where $H_{B}$ now is (\ref{eq:CH2}) expressed is the form of (\ref{reduce-h}) shown by Tolman.\footnote{See \cite{bib:tolman} eq. (102.6) for more information.}\\

@@

that corresponds to the classical $H$-functional in equilibrium described in the energy space.}\\
